Supplementary Material for: A Rough Guide to Pre-processing High-Frequency Animal Tracking Data

### 2020-11-20 _01_background.Rmd

```{r}
# do smoove data using a RACVM for three different rotational speed
data_thermals <- Map(function(rot_speed, patch) {
 dt <- smoove::simulateRACVM(dt = 0.1, Tmax = 50, omega = rot_speed, v0 = 0.1) 
 dt <- as.data.table(dt$XY)
 dt[, patch := patch]
}, c(7, 7, 7), sprintf('patch_%i', seq(3)))

# rescale x and y
data_thermals <- Map(function(df, lim1, lim2, t1, t2) {
  df[, x := x + lim1]
  df[, y := y + lim2]
  df[, time := seq(t1, t2)]
}, data_thermals, c(1, 5, 9), c(0, 2.5, 5),
   c(100, 800, 1500), c(600, 1300, 2000))

@@

# bind
gaps <- rbind(starts, ends)
setorder(gaps, "time")

# get transition phases
# THIS IS HARDCODED BECAUSE THERE DOESN'T APPEAR TO BE A BETTER WAY
# FOR THE SAME TIME
transit_1 <- gaps[c(2, 3), ]
transit_2 <- gaps[c(4, 5), ]

# simulate movement
transits <- Map(function(df){
  time_lim <- as.vector(df$time)

  track <- as.data.table(approx(df$x, df$y, n = 10))

  # add some error
  track[, `:=`(y = do_add_error(y, 0.01))]

  track[, `:=`(time = seq(time_lim[1], time_lim[2], length.out = 10),
               patch = "transit")]

  return(track)

}, list(transit_1, transit_2))

# bind
transits <- rbindlist(transits)
```

```{r}
# bind transits to thermals
data_move <- merge(data_thermals, transits, on = "time", all = T)
setorder(data_move, time)

>>>>

```{r}
# set a seed
set.seed(0)

# do smoove data using a RACVM for three different rotational speed
data_thermals <- Map(function(rot_speed, patch) {
 dt <- smoove::simulateRACVM(dt = 0.1, Tmax = 50, omega = rot_speed, v0 = 0.1) 
 dt <- as.data.table(dt$XY)
 dt[, patch := patch]
 dt[, time := seq(.N)]
}, c(7, 7, 7), sprintf('patch_%i', seq(3)))

# rescale x and y
data_thermals <- Map(function(df, lim1, lim2, offset) {
  df[, x := x + lim1]
  df[, y := y + lim2]
  df[, time := time + offset]
}, data_thermals, c(1, 5, 9), c(0, 2.5, 5),
   c(100, 800, 1500))

@@

# bind
gaps <- rbind(starts, ends)
setorder(gaps, "time")
gaps[, event := rep(c("start", "end"), 3)]

# cast
gaps <- dcast(gaps, patch ~ event, value.var = c("x", "y", "time"))

# make a trajectory section
transit <- data.table(
  x1 = shift(gaps$x_end),
  x2 = gaps$x_start,
  y1 = shift(gaps$y_end),
  y2 = gaps$y_start,
  t1 = shift(gaps$time_end),
  t2 = gaps$time_start,
  transit = sprintf("transit_%i", seq(nrow(gaps))),
  pseudo_time = seq(nrow(gaps))
)

# nafill
transit[, c("x1", "y1", "t1") := 
          lapply(list(x1, y1, t1), nafill, fill = 0)]

# interpolate
transit[, d := Map(function(x1, x2, y1, y2) {
  as.data.table(approx(x = c(x1, x2), 
                       y = c(y1, y2), 
                       n = 20))
  }, x1, x2, y1, y2)]

# add errors
transit[, d := Map(function(df) {
  df[, lapply(.SD, do_add_error, std_dev = 0.05)]
}, d)]

# get transit data
transit_data <- Map(function(df, t1, t2) {
  df <- copy(df)
  df[, time := seq(t1, t2, length.out = nrow(df))]
}, transit$d, transit$t1, transit$t2)

# merge with thermals
transit_data <- rbindlist(transit_data)
```

```{r}
# bind transits to thermals
data_move <- merge(data_thermals, transit_data, 
                   on = "time", all = T)
setorder(data_move, time)

### 2020-11-20 _02_filtering_data.Rmd

```{r remove_outside_bbox, eval=FALSE}
# remove positions outside a bounding box
# NB: set remove_inside to FALSE
data_inside_bbox <- atl_filter_bounds(data = data_copy,
                                      y_range = c(0.5, 1),
                                      remove_inside = FALSE)
```

@@

```{r example_remove_outliers, eval=FALSE}
# get speed and turning angle
data_copy[, `:=`(in_speed = atl_get_speed(data_copy,
                                          type = "in"),
                 out_speed = atl_get_speed(data_copy,
                                           type = "out"),
                 angle = atl_turning_angle(data_copy))]
```

>>>>

```{r remove_outside_bbox}
# remove positions outside a bounding box
# NB: set remove_inside to FALSE
data_inside_bbox <- atl_filter_bounds(data = data_copy,
                                      y_range = c(0.5, 1),
                                      remove_inside = FALSE)
```

@@

```{r example_remove_outliers}
# get speed and turning angle
data_copy[, `:=`(in_speed = atl_get_speed(data_copy,
                                          type = "in"),
                 out_speed = atl_get_speed(data_copy,
                                           type = "out"),
                 angle = atl_turning_angle(data_copy))]
```
